---
title: "CSS bird exclusion"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.width = 6, cache=TRUE)

setwd("/Users/colleennell/Dropbox/Projects/CSS")

library(devtools)
library(ggplot2)
library(dplyr)
library(reshape2)
library(cowplot)
library(tidyr)
library(vegan)

source('/Users/colleennell/Dropbox/rstats/themes/theme_nell.R')

se<-function(x) sd(x, na.rm=TRUE)/sqrt(length(x))

```  

```{r}

sizes<-read.csv('CSS_sizes.csv')##arthropod broken up by size class and effort sorting?
methods<-read.csv('CSS_methods.csv')
arths<-read.csv('CSS_arths.csv') #total arthropods plus measurements for complexity, bees, volume, plant wt vacuum time. 
orders<-read.csv('final/CSS_order.csv')%>%select(-contains('X'), -inits, -PLANT)

hpq<-read.csv('CSS_HPQ.csv')
hpq_sp<-read.csv('CSS_HPQ_SP.csv')

```  
Methods: These data are from a field experiment conducted at the former IRC seed farm in Irvine from Feb 2014-June 2014. For 9 plant species, a treatment of bird exclusion netting was applied to 8 experimental plants, paired with 5 control plants with no exclusion netting. After 3 months, the arthropods on the plants were sampled using vaccuum collection and identified to order or subgrouping to determine trophic feeding group. In addition, experimental plants were measured for their size and complexity, as well as host platn quality in a laboratory feeding assay of Spodoptera exigua.  

Questions:  
Does arthropod density (in exclusion) relate to HPQ? complexity?  
Does Indirect defense (LRR) related to direct defenses (HPQ & density in exclusion)? Complexity?    
Does community composition differ with bird effects? (permanova composition, exclusion vs not)  
What arthropod Orders are consumed by birds?  
Does bird feeding alter the average size of herbivores?  

To do:  
Arthropods density  
Arthropod guild/functional group by order  
Calculate LRR  
PIC  
Complexity - the number of intersections / total width OR  standard deviation of the length between intersections  

```{r}
size.melt<-sizes[-1,]%>%dplyr::select(SPECIES, Sample, treat, starts_with('X'))%>%dplyr::select(-X, -X.1, -X.2, -X.3, -X.4, -X.5, -X.6, -X.7)
sum(sizes$ABUNDANCE, na.rm=TRUE)

lg.size<-size.melt%>%select(SPECIES, Sample, treat, contains('.1'))%>%melt(id.vars=c('SPECIES','Sample','treat'))%>%mutate(size='large')
sm.size<-size.melt%>%select(-contains('.1'))%>%melt(id.vars=c('SPECIES','Sample','treat'))%>%mutate(size='small')
allsize<-rbind(lg.size, sm.size)

##total abundance of arthropods by size
all.total<-allsize%>%filter(SPECIES!='')%>%transform(value=as.numeric(value))%>%group_by(SPECIES, treat, size)%>%summarize(total = sum(value, na.rm=TRUE), mean = mean(value, na.rm=TRUE), se=se(value))

##overall
totes<-allsize%>%filter(SPECIES!='')%>%transform(value=as.numeric(value))%>%group_by(SPECIES, treat)%>%summarize(total = sum(value, na.rm=TRUE), mean = mean(value, na.rm=TRUE), se=se(value))

ggplot(totes, aes(x=SPECIES, y= mean))+geom_bar(stat='identity', aes(fill=treat), position='dodge')+geom_errorbar(aes(ymin=mean-se, ymax=mean+se), position=position_dodge(.9), width=.1)

sum(totes$total)
##recorded 2176 total arthropods collected across 116 plants

#116 plants, 8 speces

#slides say 2650 arthropods?
```


Order ID data  
116 plants, a total of 2176 arthropods  
```{r}
str(orders)

sum(orders$value, na.rm=TRUE)
##2091 arthropods- only 100 short from total counts

##add grouping for treatment
groupings<-sizes[-1,]%>%select(SPECIES, Sample, treat)%>%unique()
groupings


order.t<-full_join(orders, groupings, by='Sample')
str(order.t)
write.csv(order.t, 'CSS_arth_new.csv',row.names=FALSE)

exp.d<-order.t%>%group_by(SP, treat, Sample)%>%summarize(n=length(unique(Sample)), total_abun = sum(abundance, na.rm=TRUE), ids = paste(unique(Sample), collapse =', '))

##this one adds up as hoped
ARTMAT<-read.csv('CSS_master_MAT.csv')
ARTMAT[is.na(ARTMAT)]<-0
gps<-ARTMAT%>%group_by(species, treat)%>%summarize(length(Sample))

##community matrix
mat<-ARTMAT%>%select(ACAR:WASP)
mat.o<-ARTMAT%>%mutate(HYM = ANT+BEE+HYME+WASP, HEMI = AUCH+HETE+MIRID+STER, COL=COCC+COLE, abun=rowSums(mat, na.rm=TRUE))%>%select(-ANT, -BEE, -HYME, -AUCH, -HETE, -MIRID, -STER, -COCC, -COLE)

#calcualte total abundance, plant size, density
ath.d<-ARTMAT%>%mutate(plant_wt = WT_multiplier*WT_sample, abun=rowSums(mat, na.rm=TRUE), arth_dens=abun/plant_wt)
#relative abundances
arth.rel<-ath.d%>%select(ACAR:WASP, abun)%>%mutate_all(funs(./abun))%>%select(-abun)
arth.rel[is.na(arth.rel)]<-0
##densities
arth.dens<-ath.d%>%select(ACAR:WASP, plant_wt)%>%mutate_all(funs(./plant_wt))%>%select(-plant_wt)
arth.dens[is.na(arth.dens)]<-0
dens.dist<-vegdist(arth.dens, method='bray')

##does arthropod composition differ with bird feeding?
arth.dist<-vegdist(arth.rel, method='jaccard')
rel.comp<-adonis(arth.dist~treat*species, data=ath.d)
rel.comp

##arthropod composition differs by species, and treatment*species interaction but no main effect of treatment - this is using relative abundances
arth.mds<-metaMDS(arth.dist, trymax=100, k=3)
arth.mds
plot(arth.mds)

dens.comp<-adonis(dens.dist~treat*species, data=ath.d)
dens.comp
##similar result

dens.mds<-metaMDS(dens.dist, trymax=100, k=2)
dens.mds##do not converge

ord.rel<-mat.o%>%select(ACAR:WASP, HYM, HEMI, COL, abun)%>%mutate_all(funs(./abun))%>%select(-abun)
#ord.rel[is.na(ord.rel)]<-0
ord.dist<-vegdist(ord.rel, method='bray')
ord.comp<-adonis2(ord.dist~treat, data=ath.d, strata='species')
ord.comp
ord.mds<-monoMDS(ord.dist, trymax=100)
ord.mds
ord.df<-data.frame(ord.mds$points)%>%mutate(treat=ath.d$treat, sp = ath.d$species, dens = ath.d$arth_dens)

ggplot(ord.df, aes(x=MDS1, y=MDS2))+geom_point(aes(color=sp, shape=treat, size=dens))

a.cap<-capscale(ord.dist~Condition(species), data=ath.d, comm=arth.dens)
head(summary(a.cap))
plot(a.cap)
pdf<-data.frame(scores(a.cap)$sites)%>%mutate(sp=ath.d$species, treat=ath.d$treat)

ggplot(pdf, aes(x=MDS1, y=MDS2))+geom_point(aes(color=sp, shape=treat), size=3)
```  

Are there size differences in prey between treatments?  

## Arthropod densities by bird exclusion treatment  

```{r}
arths<-read.csv('CSS_master_MAT.csv')
str(arths)
arth.mat<-arths%>%dplyr::select(ACAR:WASP)

##calculate arthropod abundance, density, plant dry mass
arths<-arths%>%
  mutate(abun = rowSums(arth.mat, na.rm=TRUE),
         plant_g = WT_multiplier*WT_sample,
         dens = abun/plant_g)

##average arthropod density by treatment, species
arth.treat<-arths%>%
  group_by(species, treat)%>%
  summarize(dens_mean = mean(dens), dens_se = se(dens), abun_mean = mean(abun), abun_se = se(abun),
            n_plant = length(unique(Sample)))

##
```
```{r, fig.height=6, fig.width=6}
##plot arthropod densities
ggplot(arth.treat, aes(x=treat, y=dens_mean))+
  geom_point()+
  geom_errorbar(aes(ymin=dens_mean-dens_se, ymax=dens_mean+dens_se), width=.2)+
  facet_wrap(~species)+
  theme_nell()
##densities differ on...ardo, enca, same, saap?

```
```{r}
##abundance
ggplot(arth.treat, aes(x=treat, y=abun_mean))+
  geom_point()+
  geom_errorbar(aes(ymin=abun_mean-abun_se, ymax=abun_mean+abun_se), width=.2)+
  facet_wrap(~species)+
  theme_nell()
#abundance differs on ardo, enca, isme, lual, saap, same?
#similar but more variation

##use rarefaction to predict total abundance at equal sampling effort?
##vacuum times are given, so are searching
##rarefaction
##for each species, treatment
##extrapolate the total abundance based on vaccuum time
##total density
##vaccum time per plant volume?

str(methods) ##this is missing 2 plants??

##time to seconds
#methods$vac<-strptime(as.character(methods$vaccum), format = '%M:%S')
methods$vac_sec<-sapply(strsplit(as.character(methods$vaccum), ':'), 
       function(x){
         x<-as.numeric(x)
         x[1]*60+x[2]})
methods$vac_min<-methods$vac_sec/60

##sampling effort per plant
##plant volume/dry mass vaccumed per minute
methods$vac_g_min<-methods$plant_wt/methods$vac_min
methods$dens_g_min<-methods$abun/methods$vac_g_min
write.csv(methods, 'final/CSS_methods.csv', row.names=FALSE)

ggplot(methods, aes(x=vac_g_min, y=abun, color=treat))+geom_point()#+facet_wrap(~sp)
##because plants were sampled in similar effort
#plant dry wt is not related to total abundance?

```
```{r}
methods<-read.csv('final/CSS_methods.csv')
## go with composition weights and abundances
str(methods)
##bigger plants should have higher insect abundance
##insect density is related to quality?

##compare abundance to composition data
meth<-methods%>%left_join(arths%>%dplyr::select(Sample, abun_comp = abun, plant_g), by=c('Sample'))%>%
  dplyr::select(-plant_wt, -abun)%>%
  mutate(plant_kg=plant_g/1000, min_g = vac_min/plant_kg, dens_min_g = (abun_comp/min_g)/plant_kg)
write.csv(meth, 'final/CSS_methods_2017.csv', row.names=FALSE)


```
```{r}
#density by trea,t sp
dens<-meth%>%
  group_by(sp, treat)%>%
  summarize(mean = mean(dens_min_g, na.rm=TRUE), se=se(dens_min_g))

ggplot(dens, aes(x=treat, y=mean, color=sp))+
  geom_point()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2)+
  facet_wrap(~sp)

gmin_aov<-aov(dens_min_g~sp*treat, data=meth)
summary(gmin_aov)
shapiro.test(resid(gmin_aov))

## insects per g-1 min-1
# insect densities differed by species, not treatment - variation in quality
# 

##determine pred/herb densities
##add plant volume info


new<-read.csv('')

ggplot(methods)+geom_histogram(aes(x=dens_g_min))

```  
## Plant traits  


